---
title: "Mapping oil spills in California"
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library(tidyverse)
library(here)
library(janitor)
library(sf)
library(tmap)
library(expss)
```

```{r}
# read in spatial data
counties <- read_sf(here("_posts", "2021-03-14-map", "ca_counties", layer = "CA_Counties_TIGER2016.shp")) %>% 
  clean_names()

oil_spills <- read_sf(here("_posts", "2021-03-14-map","Oil_Spill_Incidents", layer = "Oil_Spill_Incident_Tracking_%5Bds394%5D.shp")) %>% 
  clean_names()

# check crs
# st_crs(counties) # WGS 84
# st_crs(oil_spills) # WGS 84

```

## Oil spill locations
```{r, layout="l-body-outset"}
# set to interactive viewing mode
tmap_mode("view")

# make exploratory map
tm_shape(counties)+
  tm_borders("black")+
  tm_shape(oil_spills)+
  tm_dots("red")
```

**Figure 1. California oil spills (2008).** Oil spills that occured in California in 2008 are shown (red points, n = 3,237). Data source: [CA DFW Oil Spill Incident Tracking](https://gis.data.ca.gov/datasets/7464e3d6f4924b50ad06e5a553d71086_0/data)


## Oil spills by county
```{r, layout="l-body-outset"}
# spatial join counties with spills and summarize number of spills per county
county_spills <- st_join(counties, oil_spills) %>% 
  group_by(name) %>% 
  summarize(spill_count = count_if("Inland", inlandmari))

# make static chloropleth map of  
ggplot()+
  geom_sf(county_spills, mapping = aes(fill = spill_count))+
  theme_void()+
  scale_fill_steps(low = "white",
                    high = "#033296",
                    n.breaks = 6,
                    name = "Number of inland\noil spills")+
  theme(legend.position = c(0.85, 0.75))
```

**Figure 2. Oil spills by county (2008).** The number of inland oil spills that occurred in each county in California in 2008 is indicated by color, with darker colors indicating more spills. Data source: [CA DFW Oil Spill Incident Tracking](https://gis.data.ca.gov/datasets/7464e3d6f4924b50ad06e5a553d71086_0/data)

