---
title: "Binomial Logistic Regression with Palmettos"
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(corrplot)
library(GGally)
library(forcats)
library(broom)
library(jtools)
library(kableExtra)
library(janitor)
library(tibble)
library(here)
```

## Differences in canopy characteristics of palmetto dates *Sabal etonia* and *Serenoa repens*

```{r, warning=FALSE, message=FALSE}
# read in data

palmetto <- read_csv(here("_posts", "2021-03-14-blr","palmetto.csv")) %>% 
  select(species, height, length, width, green_lvs) %>% 
  drop_na() %>% 
  mutate(species = as.character(species)) %>% 
  mutate(species = case_when(species == "1"~ "Serenoa repens",
                              species == "2" ~ "Sabal etonia"))

# explore correlations
# ggpairs(palmetto, ggplot2::aes(color = species))
```


```{r, warning=FALSE, message=FALSE}
# Violin plot of green leaves per tree

ggplot(palmetto, aes(x = species, y = green_lvs, fill = species))+
  geom_violin(show.legend = FALSE)+
  theme_minimal()+
  labs(x = "Species", y = "Number of green leaves")
```

**Figure 1. Green leaves on palmetto date trees by species.** *Sabal etonia* (red) tends to have fewer green leaves than *Serenoa repens* (teal).


**Key takeaway:** There is less variance in the of number of green leaves per tree for *Sabal etonia* than for *Serenoa repens*. In addition, most *Sabal etonia* in the sample have 3-5 green leaves, while 6-8 green leaves is more common for *Serenoa repens* trees.

```{r, warning=FALSE, message=FALSE}
# Density plot of canopy length

ggplot(palmetto, aes(x = length, fill = species, color = species))+
  geom_density(alpha=.4)+
  theme_minimal()+
  labs(x = "Canopy length (cm)", y = "Density", color = "Species", fill = "Species")+
  theme(legend.position = c(0.8,0.8))
```

**Figure 2. Canopy length of palmetto date trees by species.** Canopy lengths of *Sabal etonia* (red) and *Serenoa repens* (teal) are normally distributed.



**Key takeaway:** While the distributions of canopy lengths of *Sabal etonia* and *Serenoa repens* are similar, *Sabal etonia* trees tend to have slightly longer canopies than *Serenoa repens*.


```{r, warning=FALSE, message=FALSE}
# Canopy width and height scatter plot

ggplot(palmetto, aes(y = height, x = width, color = species))+
  geom_point(size = 0.2, show.legend = FALSE)+
  theme_minimal()+
  labs(y = "Canopy height (cm)", x = "Canopy width (cm)")+
  facet_wrap(~species)
```

**Figure 3. Canopy height and width of palmetto date trees by species.** Canopy width and height of *Sabal etonia* (red) and *Serenoa repens* (teal) are positively correlated.



**Key takeaway:** Distributions of canopy height and width are similar for *Sabal etonia* and *Serenoa repens*



## Binary logistic regression

**Table 1. Binary logistic regression model results.**
```{r, warning=FALSE, message=FALSE}
# binary logistic regression

palmetto_factor <- palmetto %>% 
  mutate(species = as_factor(species))

palmetto_blr <- glm(species ~ height + length + width + green_lvs, 
                            data = palmetto_factor,
                            family = "binomial")

# make table of model results

blr_tidy <- broom::tidy(palmetto_blr)

blr_tidy %>% 
  kbl(col.names = c('', 'Estimate', 'Std Error', 'Statistic', 'P value'), digits = 3) %>% 
  kable_paper("hover", full_width = F)
```

Number of green leaves is the strongest predictor variable. All predictor variables are significant.

## Evaluating the model

**Table 2. Model predictions.** 
```{r, warning=FALSE, message=FALSE}
# predict species
blr_fitted <- palmetto_blr %>% 
  broom::augment(type.predict = "response") %>% 
  mutate(predicted_species = case_when(.fitted > 0.5 ~ "Sabal etonia",
                              .fitted < 0.5  ~ "Serenoa repens")) %>% 
  mutate(success = case_when(species == predicted_species ~ 1,
                            species != predicted_species ~ 0))

# Summarize predictions
palmetto_blr_summary <- blr_fitted %>% 
  group_by(species) %>% 
  summarize(correct = sum(success),
            incorrect = n()-sum(success),
            percent_correct = sum(success)/n()*100) %>% 
  rename("Number Correct" = correct, 
         "Number Incorrect" = incorrect,
         "Percent Correct" = percent_correct)
  
# Make table
palmetto_blr_summary %>% 
  t() %>% 
  row_to_names(1) %>% 
  kable(digits = 2) %>% 
  kable_paper("hover", full_width = F)

```

The model correctly classifies over 90% of plants of both species. The model is slightly better at classifying *Sabal etonia* (92.6% correct) than *Serenoa repens* (90.8% correct).

